---
params:
  seurat: "../results/seurat/seurat_after_qc.RDS"
  pipeline: "multiome"
  integrate_datasets: true
  outcomes: ["Group"]
  sketch_cells: false
  rna_normalization_method: "log"
  species: "Homo sapiens"
  cpus: 3
  cc_avg_method: "triMean"
  cc_min_cells: 10
  cc_pathways: ["LCK", 'SELPLG']
  cc_diagrams: "chord"
  cc_subset_receiver: ["CD14 Mono", "CD16 Mono", "CD4 Naive"]
  cc_subset_sender: ["CD14 Mono", "CD16 Mono", "CD4 Naive"]
  cc_sketch: false
  cc_subset: ["CD14 Mono", "CD16 Mono"]
  
---

## Cell-cell interactome

```{r setup, warning=F, message=F, echo=knitr::is_html_output(), eval=T}

selected_packages <- c("Seurat",
                       "Signac",
                       'tidyverse',
                       'readxl',
                       'patchwork',
                       'clustree', 
                       'DT',
                       'ggrastr',
                       'presto', 
                       'SingleCellExperiment',
                       'CellChat')
#invisible(extrafont::font_import(prompt = F))
selected_font <- "Arial"
selected_font_pdf <- "ArialMT"
#extrafont::loadfonts(device = "win")
invisible(lapply(selected_packages, library, character.only = TRUE))

if(knitr::is_html_output()){
  eval_set <- TRUE
  eval_pdf <- FALSE
  echo_set <- TRUE
  message_set <- FALSE
} else {
  eval_set <- FALSE
  eval_pdf <- TRUE
  echo_set <- FALSE
  message_set <- FALSE
}

source("run_funcs.R")


knitr::opts_chunk$set(echo = echo_set,  warning = FALSE, message = FALSE, fig.align = "center", eval = params$eval, fig.width = 10, fig.height = 9 ,cache.lazy = F, results = "asis")

```

```{r echo=F}

run_gex <- FALSE
run_atac <- FALSE
run_adt <- FALSE

multimodal <- FALSE


if(params$pipeline == "atac"){
  reduction_sel <- 'umap.atac'
  if(params$integrate_datasets){
    reduction_sel <- 'umap.atac.int'
  }
}

if(params$pipeline == "rna"){
  reduction_sel <- 'umap.rna'
  if(params$integrate_datasets){
    reduction_sel <- 'umap.rna.int'
  }
}

if(params$pipeline %in% c("cite", 'multiome', 'teaseq')){
  multimodal <- TRUE
  reduction_sel <- "wnn.umap"
}

if(params$pipeline != "atac"){
  run_gex <- TRUE
}

if(params$pipeline %in% c("atac", 'multiome', 'teaseq')){
  run_atac <- TRUE
}

if(params$pipeline %in% c("cite", 'teaseq')){
  run_adt <- TRUE
}

if(params$pipeline %in% c("cite", 'teaseq', 'multiome')){
  reduction_sel <- "wnn.umap"
}

```

## Analysis Parameters
```{r echo=FALSE}

df <- params_to_df(params)
df


```


### Load Data
```{r }
seurat <- readRDS(params$seurat)
if(params$rna_normalization_method == "sct"){
  DefaultAssay(seurat) <- 'SCT'
} else {
  DefaultAssay(seurat) <- 'RNA'
}
```


## Create CellChat object
```{r }

seurat_ls <- SplitObject(seurat, split.by = params$outcomes[1])

for(i in names(seurat_ls)){
  if(!params$sketch_cells){
    seurat_ls[[i]][['RNA']]$counts <- as(object = seurat_ls[[i]][['RNA']]$counts, Class = "dgCMatrix")
    seurat_ls[[i]][['RNA']]$data <- as(object = seurat_ls[[i]][['RNA']]$data, Class = "dgCMatrix")
  } 
  
  if(!"samples" %in% colnames(seurat@meta.data)){
    seurat_ls[[i]]$samples <- seurat_ls[[i]]$SampleID
  }
}


if(params$species == "Mus musculus"){
  CellChatDB <- CellChatDB.mouse
  ppi <- PPI.mouse
} else {
  CellChatDB <- CellChatDB.human
  ppi <- PPI.human
} 

ccLS <- sapply(names(seurat_ls), simplify = F, USE.NAMES = T, function(x){
  if(params$sketch_cells){
    cc <- createCellChat(object = seurat_ls[[x]], group.by = "cell_type", assay = "sketch")
  } else {
    cc <- createCellChat(object = seurat_ls[[x]], group.by = "cell_type", assay = "RNA")
  }
  
  cc@DB <- CellChatDB
  return(cc)
})



#
rm(seurat_ls)

```

## Preprocess data for cell-cell analysis
```{r}
future::plan("multisession", workers = params$cpu)


for(i in names(ccLS)){
  ccLS[[i]] <- subsetData(ccLS[[i]]) 
  ccLS[[i]] <- identifyOverExpressedGenes(ccLS[[i]])
  ccLS[[i]] <- identifyOverExpressedInteractions(ccLS[[i]])
}

#cellChat <- projectData(cellChat, ppi)
```

## Infer cellular communication network
```{r }
for(i in names(ccLS)){
  ccLS[[i]] <- computeCommunProb(ccLS[[i]], type = params$cc_avg_method)
  ccLS[[i]] <- filterCommunication(ccLS[[i]], min.cells = params$cc_min_cells)

  ccLS[[i]] <- computeCommunProbPathway(ccLS[[i]])
  ccLS[[i]]<- aggregateNet(ccLS[[i]])

}



all_subsets <- unique(unlist(lapply(ccLS, function(x){levels(x@idents)}), use.names = F))
for(i in names(ccLS)){
  ccLS[[i]] <- liftCellChat(ccLS[[i]], group.new = all_subsets)
}




cellChat <- mergeCellChat(ccLS, add.names = names(ccLS))


compareInteractions(cellChat, show.legend = F, group = seq(1, length(names(ccLS))))

compareInteractions(cellChat, show.legend = F, group = seq(1, length(names(ccLS))), measure = "weight")


weight.max <- getMaxWeight(ccLS, attribute = c("idents","count"))

for (i in 1:length(ccLS)) {
  par(mfrow = c(1,1), xpd=TRUE)
  netVisual_circle(ccLS[[i]]@net$count, weight.scale = T, label.edge= F, edge.weight.max = weight.max[2], edge.width.max = 12, title.name = paste0("Number of interactions - ", names(ccLS)[i]))
}


all_subsets <- levels(unique(unlist(cellChat@idents, use.names = F)))

if(length(params$cc_subset) == 1){
  if(params$cc_subset == "all"){
         idents_use <- NULL
         subset_cc <- FALSE
      } else {
        idents_use <- params$cc_subset
        subset_cc <- TRUE
      }
  } else {
      idents_use <- params$cc_subset
      subset_cc <- TRUE
}


```
`r if(subset_cc){"## Subset on specific cell types"}`
```{r eval=subset_cc, echo=subset_cc & echo_set}
#par(mfrow = c(length(ccLS),1), xpd=TRUE)


for (i in 1:length(ccLS)) {
  par(mfrow = c(1,1), xpd=TRUE)
  netVisual_circle(ccLS[[i]]@net$count, 
                   weight.scale = T, 
                   label.edge= F, 
                   edge.weight.max = weight.max[2], 
                   edge.width.max = 12, 
                   idents.use = params$cc_subset,
                   title.name = paste0("Number of interactions - ", names(ccLS)[i]))
}

```

## Compare Cell types in 2D space
This analysis allows more easily to identify cell types with significant changes across conditions. 

```{r }


for(i in 1:length(ccLS)){
  ccLS[[i]] <- netAnalysis_computeCentrality(ccLS[[i]])
}

num.link <- sapply(ccLS, function(x) {rowSums(x@net$count) + colSums(x@net$count)-diag(x@net$count)})
weight.MinMax <- c(min(unlist(num.link)), max(unlist(num.link))) 

for (i in 1:length(ccLS)) {
  
  p <- netAnalysis_signalingRole_scatter(ccLS[[i]], title = names(ccLS)[i], weight.MinMax = weight.MinMax)
  print(p)
}


```


## Cell-cell network visualization
### Pathway level interactions
#### Explore all pathways
```{r }

#
cellChat <- computeNetSimilarityPairwise(cellChat, type = "functional")

pathway_union <- unique(unlist(lapply(names(ccLS), function(x){ ccLS[[x]]@netP$pathways }), use.names = F))

lapply(names(ccLS), function(x){
  netAnalysis_signalingRole_heatmap(ccLS[[x]], pattern = "outgoing", signaling = pathway_union, title = x, width = 9, height = 6)
})

lapply(names(ccLS), function(x){
  netAnalysis_signalingRole_heatmap(ccLS[[x]], pattern = "incoming", signaling = pathway_union, title = x, width = 9, height = 6)
})

lapply(names(ccLS), function(x){
  netAnalysis_signalingRole_heatmap(ccLS[[x]], pattern = "all", signaling = pathway_union, title = x, width = 9, height = 6)
})




```

#### Pathways of interest
```{r }

pathways.show <- params$cc_pathways

# 
# get_connected_pathways <- function(cellChat, idents = NULL) {
#   # Ensure netP$prob exists
#   if (!"prob" %in% names(cellChat@netP)) {
#     stop("No probability matrix found in cellChat@netP. Ensure netAnalysis_computeCentrality was run.")
#   }
# 
#   # Extract the probability matrix
#   prob_matrix <- cellChat@netP$prob  # 3D array: rows (senders) x columns (receivers) x pathways
# 
#   # Get dimension names
#   sender_cells <- dimnames(prob_matrix)[[1]]
#   receiver_cells <- dimnames(prob_matrix)[[2]]
#   pathway_names <- dimnames(prob_matrix)[[3]]
# 
#   # Initialize list to store results
#   results <- list()
# 
#   # Iterate through pathways
#   for (pathway in pathway_names) {
#     matrix_slice <- prob_matrix[, , pathway]  # Extract 2D matrix for the pathway
# 
#     # Find nonzero interactions (senders and receivers)
#     sender_receiver_pairs <- which(matrix_slice > 0, arr.ind = TRUE)
# 
#     if (nrow(sender_receiver_pairs) > 0) {
#       sender_idents <- sender_cells[sender_receiver_pairs[, 1]]
#       receiver_idents <- receiver_cells[sender_receiver_pairs[, 2]]
# 
#       # Create a temporary data frame
#       pathway_df <- data.frame(
#         pathway = rep(pathway, length(sender_idents)),
#         sender_ident = sender_idents,
#         receiver_ident = receiver_idents,
#         stringsAsFactors = FALSE
#       )
# 
#       # Append to results list
#       results[[pathway]] <- pathway_df
#     }
#   }
# 
#   # Combine all results into one data frame
#   result_df <- do.call(rbind, results)
# 
#   # Apply filtering if idents is provided
#   if (!is.null(idents)) {
#     result_df <- result_df[result_df$sender_ident %in% idents | result_df$receiver_ident %in% idents, ]
#   }
# 
#   return(result_df)
# }


# 
# all_pathway <- bind_rows(lapply(names(ccLS), function(x){
#   if(subset_cc){
#     pw <- get_connected_pathways(ccLS[[x]], idents = params$cc_subset)
#   } else {
#     pw <- get_connected_pathways(ccLS[[x]], idents = params$cc_subset)
#   }
#   df <- pw %>%
#         mutate(Condition = x)
# })) %>%
#   group_by(pathway) %>%
#   mutate(n_conditions = n_distinct(Condition),
#          Detected_in = paste(unique(Condition), collapse = "; "),
#          Sender = paste(unique(sender_ident), collapse = '; '),
#          Receiver = paste(unique(receiver_ident), collapse = '; ')) %>%
#   ungroup() %>%
#   dplyr::select(-Condition, -sender_ident, -receiver_ident) %>%
#   unique() %>%
#   dplyr::arrange(desc(n_conditions))






for(i in pathways.show){
  weight.max <- getMaxWeight(ccLS, slot.name = c("netP"), attribute = i)
  for(j in names(ccLS)){
    par(mfrow = c(1,1), xpd=TRUE)
    netVisual_aggregate(ccLS[[j]], 
                        signaling = i, 
                        layout = "circle", 
                        edge.weight.max = weight.max[1], 
                        edge.width.max = 10, 
                        signaling.name = paste(i, j),
                        idents.use = idents_use,
                        remove.isolate = F)
  
  }
}

```

### Ligand-receptor pair communication probability per pathway
```{r }
comm_sub_tmp <- bind_rows(lapply(names(ccLS), function(x){
    df <- subsetCommunication(ccLS[[x]]) %>%
        mutate(dataset = x)
  }))


if(subset_cc){
    comm_sub <- comm_sub_tmp %>%
              dplyr::filter(source %in% params$cc_subset | target %in% params$cc_subset)
} else {
  comm_sub <- comm_sub_tmp
}



if(knitr::is_html_output()){
  datatable(comm_sub,
                        caption = "Ligand-receptor pair activity per pathway",
                        extensions = 'Buttons',
                          options = list(dom = 'lfrtpB',
                                         buttons = list(
                                                      list(extend = 'csv', title = "ligand_receptor_pairs"),
                                                      list(extend = 'excel', title = "ligand_receptor_pairs")),
                                         lengthMenu = list(c(10,25,50,-1),
                                                           c(10,25,50,"All")))) %>%
                    formatSignif(columns = c("pval"), digits = 3 ) %>%
                    formatRound(columns = c("prob"), digits = 3)
  
} 

lapply(intersect(pathways.show, comm_sub$pathway_name), function(x){
  comm_sub_path <- comm_sub %>%
                    dplyr::filter(pathway_name == x) 
 
  if(dim(comm_sub_path)[1] > 0){
    if(length(unique(comm_sub_path$dataset)) > 1){
      p1 <- netVisual_bubble(cellChat,
                       pairLR.use = comm_sub_path[c("interaction_name")], 
                       angle.x = 45, 
                       comparison = c(1, 2), 
                       max.dataset = 2,
                       remove.isolate = T,
                       title.name = paste("Increased signaling in", names(ccLS)[1], "vs", names(ccLS)[2] ))
      p2 <- netVisual_bubble(cellChat,
                       pairLR.use = comm_sub_path[c("interaction_name")], 
                       angle.x = 45, 
                       comparison = c(1, 2), 
                       max.dataset = 1,
                       remove.isolate = T,
                       title.name = paste("Decreased signaling in", names(ccLS)[1], "vs", names(ccLS)[2] ))
      p1 + p2
    } else {
      dataset_select <- unique(comm_sub_path$dataset)
      netVisual_bubble(ccLS[[dataset_select]], pairLR.use = comm_sub_path[c("interaction_name")], angle.x = 45, title.name = dataset_select )
    }
  }
})


netVisual_bubble(cellChat,
                       signaling = "LCK",
                       angle.x = 45, 
                       comparison = c(1, 2), 
                       #max.dataset = 2,
                       remove.isolate = T,
                       title.name = paste("Increased signaling in", names(ccLS)[1], "vs", names(ccLS)[2] ))
```

### Ligand-receptor pair Differential expression
```{r }
pos.dataset = names(ccLS)[2]
features.name = paste0(pos.dataset, ".merged")
shared_idents <- Reduce(intersect,lapply(ccLS, function(x){
  unique(x@meta$cell_type)
}))

cellChat <- identifyOverExpressedGenes(cellChat, 
                                       group.dataset = "datasets", 
                                       pos.dataset = pos.dataset, 
                                       features.name = features.name, 
                                       idents.use = shared_idents,
                                       only.pos = FALSE, 
                                       thresh.pc = 0.1, 
                                       thresh.fc = 0.05,
                                       thresh.p = 0.05, 
                                       group.DE.combined = FALSE) 


net <- netMappingDEG(cellChat, features.name = features.name, variable.all = TRUE)



lapply(pathways.show, function(x){
  comm_sub_filt <- comm_sub[comm_sub$pathway_name == x,]
  feats <- unique(unlist(str_split(unique(comm_sub_filt$interaction_name), pattern = "_", simplify = F)))
  feats <- intersect(feats, row.names(seurat))
  #print(feats)
  lapply(feats, function(y){
    VlnPlot(seurat, features = y, split.by = params$outcomes, ncol =  1, group.by = "cell_type", pt.size = 0) + ggtitle(paste(y, "-", x, "pathway"))
  })
  #plotGeneExpression(cellChat, signaling = x, split.by = "datasets", colors.ggplot = T, type = "violin")
})

#



```


### Save CellChat object
```{r save}
cellChat <- addReduction(cellChat, dr = seurat@reductions[[reduction_sel]], dr.name = "umap")
saveRDS(cellChat, "cellchat_object.RDS")
saveRDS(ccLS, "cellchat_list.RDS")
```