---
params:
  seurat: "../results/seurat/seurat_annotated.RDS"
  pipeline: "multiome"
  integrate_datasets: true
  species: "Homo sapiens"
  cpus: 2
---

## Transcription Factor Activity Computation

```{r setup, warning=F, message=F, echo=knitr::is_html_output(), eval=T}

selected_packages <- c("Seurat",
                       "Signac",
                       'tidyverse',
                       'readxl',
                       'patchwork',
                       'clustree', 
                       'DT',
                       'ggrastr',
                       'presto', 
                       'SingleCellExperiment',
                       'TFBSTools',
                       "chromVARmotifs",
                       'chromVAR',
                       'motifmatchr')
#invisible(extrafont::font_import(prompt = F))
selected_font <- "Arial"
selected_font_pdf <- "ArialMT"
#extrafont::loadfonts(device = "win")
invisible(lapply(selected_packages, library, character.only = TRUE))

if(knitr::is_html_output()){
  eval_set <- TRUE
  eval_pdf <- FALSE
  echo_set <- TRUE
  message_set <- FALSE
} else {
  eval_set <- FALSE
  eval_pdf <- TRUE
  echo_set <- FALSE
  message_set <- FALSE
}

source("run_funcs.R")


knitr::opts_chunk$set(echo = echo_set,  warning = FALSE, message = FALSE, fig.align = "center", eval = params$eval, fig.width = 10, fig.height = 9 ,cache.lazy = F, results = "asis")

```

```{r echo=F}

run_gex <- FALSE
run_atac <- FALSE
run_adt <- FALSE

multimodal <- FALSE


if(params$pipeline == "atac"){
  reduction_sel <- 'umap.atac'
  if(params$integrate_datasets){
    reduction_sel <- 'umap.atac.int'
  }
}


if(params$pipeline %in% c("cite", 'multiome', 'teaseq')){
  multimodal <- TRUE
  reduction_sel <- "wnn.umap"
}

if(params$pipeline != "atac"){
  run_gex <- TRUE
}

if(params$pipeline %in% c("atac", 'multiome', 'teaseq')){
  run_atac <- TRUE
}

if(params$pipeline %in% c('teaseq', 'multiome')){
  reduction_sel <- "wnn.umap"
}

```

## Analysis Parameters
```{r echo=FALSE}

df <- params_to_df(params)
df


```


### Load Data
```{r }
seurat <- readRDS(params$seurat)

DefaultAssay(seurat) <- 'ATAC'

```


## Load motifs
Motif matrices, stored in PFM format (or PWM), are derived from prior ChIP-Seq experiments across tissues or species. Other technologies used for the generation of PFMs are Protein-binding microarrays, SELEX, etc. Together, they are the basis of establishing which set of sequences are bound by a given transcription factor. 

Databases like [CISBP](https://cisbp.ccbr.utoronto.ca/), or [JASPAR](https://jaspar.elixir.no/), store these PFMs, which can be leveraged to scan chromatin for enrichment and activity.

```{r echo= echo_set }
if(params$species == "Mus musculus"){
  data("mouse_pwms_v2")
  pfm <- mouse_pwms_v2
  BiocManager::install("BSgenome.Mmusculus.UCSC.mm10", update = F)
  library(BSgenome.Mmusculus.UCSC.mm10)
  bsg <- BSgenome.Mmusculus.UCSC.mm10
  
} else {
  data("human_pwms_v2")
  pfm <- human_pwms_v2
  if(params$species == "Homo sapiens"){
    BiocManager::install("BSgenome.Hsapiens.UCSC.hg38", update = F)
    library(BSgenome.Hsapiens.UCSC.hg38)
    bsg <-  BSgenome.Hsapiens.UCSC.hg38
  } else if(params$species == "Macaca mulatta"){
    BiocManager::install("BSgenome.Mmulatta.UCSC.rheMac10", update = F)
    library(BSgenome.Mmulatta.UCSC.rheMac10)
    bsg <-  BSgenome.Mmulatta.UCSC.rheMac10

  }
}

```


```{r, message=F, warning=F, echo=echo_set}

seqlevelsStyle(Annotation(seurat)) <- "UCSC"
main.chroms <- standardChromosomes(bsg)
keep.peaks <- row.names(seurat)[as.logical(seqnames(granges(seurat)) %in% main.chroms)]
seurat[["ATAC"]] <- subset(seurat[["ATAC"]], features = keep.peaks )

motif.matrix <- CreateMotifMatrix(
  features = granges(seurat),
  pwm = pfm,
  genome = bsg
)

motif_obj <- CreateMotifObject(pwm = pfm, 
                               data = motif.matrix)
Motifs(seurat) <- motif_obj

```


## chromVAR
[chromVAR](https://pubmed.ncbi.nlm.nih.gov/28825706/) is a tool for inferring transcription factor-associated accessibility from peaks. It calculates an enrichment of TF activity by calculating gain or loss of accessibility within peaks sharing the same motifs while controlling for technical biases. 

```{r eval=multimodal & params$sketch_cells & params$pipeline %in% c("cite","teaseq"), echo=multimodal & echo_set & params$sketch_cells}

library(BiocParallel)

if(params$cpus <= 3){
  register(MulticoreParam(params$cpus))
} else {
  register(MulticoreParam(params$cpus -1))
}



seurat <- RunChromVAR(seurat, 
                      genome = bsg,
                      assay = "ATAC",
                      verbose = T)


```

### Cluster TF markers
```{r }
tf_markers <- FindAllMarkers(
  object = seurat,
  only.pos = TRUE,
  mean.fxn = rowMeans,
  fc.name = "avg_diff"
) 


tf_df <- convert_chromvar_features(seurat, features = unique(tf_markers$gene)) %>%
          dplyr::select(-id) %>%
          dplyr::rename(gene = "feature_name")

tf_markers <- tf_markers %>%
              left_join(tf_df, by = c("gene" = "gene")) %>%
              dplyr::rename(motif = "gene",
                            transcription_factor = "name") %>%
              dplyr::select(transcription_factor, everything())


top_tf_markers <- tf_markers %>%
                  group_by(cluster) %>%
                  slice_max(., order_by = avg_diff, n = 10) %>%
                  ungroup() 



if(knitr::is_html_output()){
  datatable(tf_markers,
                        caption = "TF activity cluster markers",
                        extensions = 'Buttons',
                          options = list(dom = 'lfrtpB',
                                         buttons = list(
                                                      list(extend = 'csv', title = "chromvar_cluster_markers"),
                                                      list(extend = 'excel', title = "chromvar_cluster_markers")),
                                         lengthMenu = list(c(10,25,50,-1),
                                                           c(10,25,50,"All")))) %>%
                    formatSignif(columns = c("p_val", "p_val_adj"), digits = 3 ) %>%
                    formatRound(columns = c("avg_diff", "pct.1", "pct.2"), digits = 3)
  
  
} else {
  knitr::kable(top_tf_markers)
}



```

### Cell type TF markers
```{r }
Idents(seurat) <- "cell_type"

tf_markers_cell <- FindAllMarkers(
  object = seurat,
  only.pos = TRUE,
  mean.fxn = rowMeans,
  fc.name = "avg_diff"
) 


tf_df_cell <- convert_chromvar_features(seurat, features = unique(tf_markers_cell$gene)) %>%
          dplyr::select(-id) %>%
          dplyr::rename(gene = "feature_name")

tf_markers_cell <- tf_markers_cell %>%
              left_join(tf_df_cell, by = c("gene" = "gene")) %>%
              dplyr::rename(motif = "gene",
                            transcription_factor = "name") %>%
              dplyr::select(transcription_factor, everything()) %>%
              dplyr::rename(cell_type = "cluster")


top_tf_markers_cell <- tf_markers_cell %>%
                  group_by(cell_type) %>%
                  slice_max(., order_by = avg_diff, n = 10) %>%
                  ungroup() 



if(knitr::is_html_output()){
  datatable(tf_markers_cell,
                        caption = "TF activity cell markers",
                        extensions = 'Buttons',
                          options = list(dom = 'lfrtpB',
                                         buttons = list(
                                                      list(extend = 'csv', title = "chromvar_cell_markers"),
                                                      list(extend = 'excel', title = "chromvar_cell_markers")),
                                         lengthMenu = list(c(10,25,50,-1),
                                                           c(10,25,50,"All")))) %>%
                    formatSignif(columns = c("p_val", "p_val_adj"), digits = 3 ) %>%
                    formatRound(columns = c("avg_diff", "pct.1", "pct.2"), digits = 3)
  
  
} else {
  knitr::kable(top_tf_markers_cell)
}

Idents(seurat) <- "seurat_clusters"

```

```{r }

DefaultAssay(seurat) <- "chromvar"
p <- FeaturePlot(seurat, features = ConvertMotifID(seurat[['chromvar']], name = c("CREB1", "FOXP3")))

p <- FeaturePlot(seurat, 
                 features =  convert_chromvar_features(seurat, features = "POU2F2")$feature_name,
                 min.cutoff = 'q10', 
                 max.cutoff = 'q90',
                 reduction = reduction_sel,
                 cols = c("lightgrey", "darkgreen"),
                 pt.size = 0.1)



```



```{r }
saveRDS(seurat, file = "seurat_chromvar.RDS")

```